---
title: "Using phantasusLite package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using phantasusLite package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`phantasusLite` is an R-package to integrate the work with RNA-Seq count matrices into a single and fast R-pipeline. This R-package supports loading the data from GEO by GSE. It provides url access to the remote repository with the archs4, archs4_zoo and dee2 HDF5-files for getting the count matrix. Finally `phantasusLite` allows to get an ExpressionSet with the expression matrix for future differential expression analysis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation

To install the package you should:

1.  Install R-package rhdf5client from the github
2.  Install R-package phantasusLite from the github

```{r}
library(devtools)
install_github("assaron/rhdf5client")
install_github("rsablina/phantasusLite")
```

## Loading necessary libraries

```{r message=FALSE, warning=FALSE}
library(GEOquery)
library(phantasusLite)
```

## Quick start

To run the package enter the code sample below.

Let's load the ExpressionSet from GEO

```{r message=FALSE}
ess <- getGEO("GSE85653")
es <- ess[[1]]
```

#### Working with the expression matrices

ExpressionSet from the GEO doesn't contain the expression matrix -- exprs(es) is empty.

```{r}
head(exprs(es))
```

Expression matrices are stored in HDF5-files in the remote repository. The remote repository URL is '<https://ctlab.itmo.ru/hsds/?domain=/counts>'.

```{r}
url <- 'https://ctlab.itmo.ru/hsds/?domain=/counts'
```

To get the list of all HSDS-collections in the repository:

```{r}
getHSDSCollectionsList(url)
```

To get the list of all HDF5-files on the server:

```{r}
getHSDSFileList(url)
```

To get the list of HDF5-files in the specific collection:

```{r}
collection <- 'archs4'
getHSDSCollectionFileList(url, collection)
```

One can get an expression matrix from the specific HDF5-file:

```{r}
#file <- "/counts/archs4/mouse_matrix_v11.h5"
#es1 <- loadCountsFromH5FileHSDS(es, url, file)
```

One can get an expression matrix from the HDF5-file with the data for the biggest number of samples.

```{r}
es2 <- loadCountsFromHSDS(es, url)
```

Now ExpressionSet contains an expression matrix for the samples:

```{r}
head(exprs(es))
```

#### Working with the condition
